Neonatal models

### 2021-12-01 getdata.Rmd
correct data

```{r readobs}
bb <- obs$id[obs$starttyp == 2]
births <- obs[obs$id %in% bb, 
    c("id", "sex", "birthdate", "nofrs", "sluttyp", "socStatus", "hisclass", 
      "enter", "exit", "ort", "ortnmn")]
births$event <- as.numeric(births$sluttyp == 2)
births <- age.window(births, c(0, 1))
births <- births[order(births$id, births$enter), ]
##indx <- tapply(births$id, births$id)
##births$mexit <- tapply(births$exit, births$id, max)[indx]
##births$mevent <- tapply(births$event, births$id, max)[indx]
##births <- births[!duplicated(births$id), ]
### Deaths:
did <- unique(births$id[births$event == 1])
deaths <- births[births$id %in% did, ]
indx <- tapply(deaths$id, deaths$id)
deaths$menter <- tapply(deaths$enter, deaths$id, min)[indx]
deaths <- deaths[!duplicated(deaths$id), ]
deaths$enter <- deaths$menter
deaths$menter <- NULL
deaths$enter <- 0 ## satisfied....
### Survivors:
survs <- births[!(births$id %in% did), ]
indx <- tapply(survs$id, survs$id)
survs$mexit <- tapply(survs$exit, survs$id, max)[indx]
survs <- survs[!duplicated(survs$id), ]
survs$exit <- survs$mexit
survs$mexit <- NULL
survs$enter <- 0 ## Ok ...
### Now, give birth to NEW births:
births <- rbind(deaths, survs)
births <- births[births$birthdate > 1895, ]

@@

```{r makecommunal}
##source("R/make_comm.R")
lagg <- 3 / 365  ## NOTE!!
## Ume:
bume <- births[births$subreg == "ume", ]
bume <- make.communal(bume, umetemp["heat"], start = 1895, period = 1/52, lag = lagg)
bume <- bume[bume$enter < bume$exit, ]
bume <- make.communal(bume, umetemp["cold"], start = 1895, period = 1 / 52, lag = lagg)
bume <- bume[bume$enter < bume$exit, ]
bume <- make.communal(bume, umetemp["week"], start = 1895, period = 1 / 52, lag = lagg)
bume <- bume[bume$enter < bume$exit, ]
### Ske:
bskel <- births[births$subreg == "bjur", ]
bskel <- make.communal(bskel, bjurtemp["heat"], start = 1895, period = 1/52, lag = lagg)
bskel <- bskel[bskel$enter < bskel$exit, ]
bskel <- make.communal(bskel, bjurtemp["cold"], start = 1895, period = 1 / 52, lag = lagg)
bskel <- bskel[bskel$enter < bskel$exit, ]
bskel <- make.communal(bskel, bjurtemp["week"], start = 1895, period = 1 / 52, lag = lagg)
bskel <- bskel[bskel$enter < bskel$exit, ]
##bsten <- births[births$subreg == "sten", ]
##bsten <- make_comm(bsten, stentemp, start = 1899, period = 1/52, lag = lagg)
bume$region <- "ume"
bskel$region <- "ske"
infdat <- rbind(bume, bskel)
infdat <- infdat[order(infdat$id, infdat$enter), c("region", "id", "sex", "enter", "exit", "event", 
                                                   "birthdate", "socStatus", "hisclass", "illeg",
                                                   "parity", "heat", "cold", "week")]
saveRDS(infdat, file = "data/infdat.rds")
```

>>>>

```{r readobs}
bb <- obs$id[obs$starttyp == 2]
births <- obs[obs$id %in% bb, 
    c("id", "sex", "birthdate", "nofrs", "sluttyp", "socStatus", "hisclass", 
      "enter", "exit", "ort", "ortnmn")]
births$event <- as.numeric(births$sluttyp == 2)
## NEW (and necessary):
extend <- with(births, (exit <= enter) & (event == 1))
births$exit[extend] <- births$enter[extend] + 1 / (4 * 365)
## NOW, remove all zero-length intervals:
births <- births[births$exit > births$enter, ]
## End NEW, now OK to select:
births <- age.window(births, c(0, 1))
births <- births[order(births$id, births$enter), ]
##indx <- tapply(births$id, births$id)
##births$mexit <- tapply(births$exit, births$id, max)[indx]
##births$mevent <- tapply(births$event, births$id, max)[indx]
##births <- births[!duplicated(births$id), ]
### Deaths:
did <- unique(births$id[births$event == 1])
deaths <- births[births$id %in% did, ]
indx <- tapply(deaths$id, deaths$id)
deaths$menter <- tapply(deaths$enter, deaths$id, min)[indx]
deaths <- deaths[!duplicated(deaths$id), ]
deaths$enter <- deaths$menter
deaths$menter <- NULL
deaths$enter <- 0 ## satisfied....
### Survivors:
survs <- births[!(births$id %in% did), ]
indx <- tapply(survs$id, survs$id)
survs$mexit <- tapply(survs$exit, survs$id, max)[indx]
survs <- survs[!duplicated(survs$id), ]
survs$exit <- survs$mexit
survs$mexit <- NULL
survs$enter <- 0 ## Ok ...
### Now, give birth to NEW births:
births <- rbind(deaths, survs)
births <- births[births$birthdate > 1895, ]

@@

```{r makecommunal}
##source("R/make_comm.R")
lagg <- 3 / 365  ## NOTE!!
##lagg <- 0
## Ume:
bume <- births[births$subreg == "ume", ]
bume <- make.communal(bume, umetemp["heat"], start = 1895, period = 1/52, lag = lagg)
bume <- bume[bume$enter < bume$exit, ]
bume <- make.communal(bume, umetemp["cold"], start = 1895, period = 1 / 52, lag = lagg)
bume <- bume[bume$enter < bume$exit, ]
bume <- make.communal(bume, umetemp["week"], start = 1895, period = 1 / 52, lag = lagg)
bume <- bume[bume$enter < bume$exit, ]
### Ske:
bskel <- births[births$subreg == "bjur", ]
bskel <- make.communal(bskel, bjurtemp["heat"], start = 1895, period = 1/52, lag = lagg)
bskel <- bskel[bskel$enter < bskel$exit, ]
bskel <- make.communal(bskel, bjurtemp["cold"], start = 1895, period = 1 / 52, lag = lagg)
bskel <- bskel[bskel$enter < bskel$exit, ]
bskel <- make.communal(bskel, bjurtemp["week"], start = 1895, period = 1 / 52, lag = lagg)
bskel <- bskel[bskel$enter < bskel$exit, ]
##bsten <- births[births$subreg == "sten", ]
##bsten <- make_comm(bsten, stentemp, start = 1899, period = 1/52, lag = lagg)
bume$region <- "ume"
bskel$region <- "ske"
infdat <- rbind(bume, bskel)
infdat <- infdat[order(infdat$id, infdat$enter), c("region", "id", "sex", "enter", "exit", "event", 
                                                   "birthdate", "socStatus", "hisclass", "illeg",
                                                   "parity", "heat", "cold", "week")]
saveRDS(infdat, file = "data/infdat.rds")
```

### 2021-12-01 neo.Rmd
correct data

```{r tpchreg_g, results = 'asis'}
(system.time(fit.pch_g <- tpchreg(oe(event, exposure) ~ 
                                    sex + socStatus + illeg + parity + cold,
                                time = age, data = tgneo)))
dr2_g <- drop1(fit.pch_g, test = "Chisq")
ltx(fit.pch_g, dr = dr2_g, label = "tab:tpchreg_g", 
    caption = "Piecewise constant proportional hazards model with time transformation.")
```

Compare the Tables \@ref(tab:tpchreg) and \@ref(tab:tpchreg_g). Identical except for *TTR*
(Total Time at Risk). And graphs will be stretcehd.

```{r haz_g, fig.cap = "Baseline hazard function, transformed time.",  fig.height = 4}
par(las = 1)
plot(fit.pch_g, fn = "haz")
abline(h = 0)
```

>>>>

```{r tpchreg.g, results = 'asis'}
(system.time(fit.pch.g <- tpchreg(oe(event, exposure) ~ 
                                    sex + socStatus + illeg + parity + cold,
                                time = age, data = tgneo)))
dr2.g <- drop1(fit.pch.g, test = "Chisq")
ltx(fit.pch.g, dr = dr2.g, label = "tab:tpchreg.g", 
    caption = "Piecewise constant proportional hazards model with time transformation.")
```

Compare the Tables \@ref(tab:tpchreg) and \@ref(tab:tpchreg.g). Identical except for *TTR*
(Total Time at Risk). And graphs will be stretced.

```{r haz.g, fig.cap = "Baseline hazard function, transformed time.",  fig.height = 4}
par(las = 1)
plot(fit.pch.g, fn = "haz")
abline(h = 0)
```
